---
title: "Customizing the Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Customizing the Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Although IBCS recommend to use gray colors for plotting, `tidycharts` enables user to define their color palette. The same applies to styles, the default IBSC are implemented, but there is a way to set one's own styles. In this vignette we will show how to customize the package. 

```{r setup, message=FALSE}
library(tidycharts)
```

## Colors

### The defaults

```{r defaults}
data_time_series <- data.frame(
  time = month.abb[1:8],
  Poland = round(2 + 0.5 * sin(1:8), 2),
  Germany = round(3 + sin(3:10), 2),
  Slovakia = round(2 + 2 * cos(1:8), 2)
)
column_chart(data_time_series, x = 'time', 
             series = c('Poland', 'Germany', 'Slovakia'), interval = 'months') %>% SVGrenderer()
```

### Customization

Firstly, we need to define a data frame with 6 rows and 2 columns. Column `bar_colors` contains 6 colors which will be used to draw the bars. Column `text_colors` contains color of labels that will be drawn on bars. Ideally, `text_colors` are contrasting to corresponding row of `bar_colors`. Users with a lack of the sense of aesthetics are encouraged to select color palette with a help of dedicated software :

- [colorbrewer](https://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3)
- [coolors.co](https://coolors.co/)

```{r colors}
color_df <- data.frame(
  bar_colors = c("rgb(61, 56, 124)", 
                 "rgb(0,89,161)", 
                 "rgb(0,120,186)", 
                 "rgb(0,150,193)",
                 "rgb(0, 178, 184)", 
                 "rgb(0,178,184)"),
  text_colors = c("white", 
                  "white", 
                  "white", 
                  "white",
                  "white", 
                  "black")
)
```

Then, we use `set_colors` function from `tidycharts` package.

```{r set-colors}
set_colors(color_df)
```

Now, we can generate chart using new palette. 

```{r customized-chart}
column_chart(data_time_series, x = 'time', 
             series = c('Poland', 'Germany', 'Slovakia'), interval = 'months') %>% SVGrenderer()
```

## Styles

### The defaults
Styles are used to indicate different type of presented data. `actual`, `previous`, `plan`, `forecast` styles are implemented out of the box. 

```{r styles-default}
column_chart(data_time_series, x = 'time', 
             series = 'Poland', interval = 'months',
             styles = c(rep('actual',7), 'plan')) %>% SVGrenderer()
```

### Customization

When defining styles, one must define `stroke` and `fill` parameters as column names. Style names are unrestricted. Use `set_styles` to bind `styles_df` to the package environment.  

```{r user-styles}
styles_df <- rbind(
  actual = list(fill = "rgb(64,64,64)", stroke = "rgb(64,64,64)" ),
  fictual = list(fill = "rgb(221,28,119)", stroke = "rgb(136,86,167)")
)

set_styles(styles_df)
```

Now, refer to styles through given names and use them in chart functions.

```{r styles-custom}
column_chart(data_time_series, x = 'time', 
             series = 'Poland', interval = 'months',
             styles = c(rep('actual',7), 'fictual')) %>% SVGrenderer()
```

## Resetting to defaults

If you want to restore the default styles and colors, just use `restore_defaults` function.

```{r}
restore_defaults()
```

