---
title: "Getting_Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting_Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows the usage of the ibcsplot package. It contains different chart types examples and tips for proper data visualization.

## Prerequisites
This package and vignette are created for a user who:

* has basic programming skills in R
* is familiar with R data structures, such as vector, data.frame

# How to download and install the package?

```{r setup, message=FALSE}

# install from CRAN
# install.packages(ibcsplot)

library(ibcsplot)

```

# Bar charts

## Basic

Bar charts shoud be used to show structure in one moment of time. One of typical usecase of the barchart is to visualize profit of a company in a division by departments. The data structure could look the following:

```{r data}
data_barchart <- data.frame(
  dep = c('Services', 'Production', 'Marketing', 'Purchasing'),
  profit = c(12, 15, 2, -3)
)
```

Creation of the barchart is simple. We use `barchart_plot` function to do that. After calling the function, `SVGrenderer()` should be used to render the plot as HTML content. Otherwise one element character vector with SVG content is returned. We can chain the commands by pipe operator (`%>%`).

```{r base-barchart}

barchart_plot(data = data_barchart, cat = data_barchart$dep, series = 'profit') %>% 
  SVGrenderer()

```

A plot should contain an informative title. We can use `add_title` function to make one. As with `SVGRenderer`, we chain the commands. 

```{r barchart-title}
barchart_plot(data = data_barchart, cat = data_barchart$dep, series = 'profit') %>% 
  add_title('The company XYZ', 'Profit', 'in mEUR', 'by departments, 2020') %>% 
  SVGrenderer()
```

## Indexed

We use indecies to show a reference value on the plot. In the following example, index line is used to show the best result in previous year (PY). 

```{r barchart-index}
barchart_plot_index(data = data_barchart,
                    cat = data_barchart$dep, series = 'profit', 
                    index_val = 10, index_label = 'PY best result') %>% 
  SVGrenderer()
```

